Supporting Information for ``The dynamics of starvation and recovery''

### 2017-11-03 NatureComm_REVISION/YKR_supp-Revision-1.tex
revision 3 figure changes

We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
\begin{align}
\begin{split}
\dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
& \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
\end{split}
\end{align}
where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text). In the above equations $Y$ represents the yield coefficient (e.g., Refs. \citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism). If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$. This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density. In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)

@@

We can formally non-dimensionalize this system by choosing the general rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
\begin{align}
\begin{split}
\dot{F} &= \frac{1}{s}\left[\lambda_{max} F + \rho_{max}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
\dot{H} &= \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{max}\frac{R}{qk} H - \mu H\right],  \\
\dot{R} &= \frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right) F\right]\right].
\end{split}
\end{align}
If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
\begin{align}
\begin{split}
\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
\end{split}
\end{align}
where we have dropped the subscripts on $\lambda_{max}$ and $\rho_{max}$ for simplicity, and $\xi=C/k$, $\delta=Y_{H}kP_{H}/C$, and $\beta=Y_{H}k\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)/C$. The above equations represent the system of equations presented in the main text.
\\

@@

>>>>

We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
\begin{align}
\begin{split}
\dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
& \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
\label{bigdynamics}
\end{split}
\end{align}
where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text). In the above equations $Y$ represents the yield coefficient (e.g., Refs. \citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism). If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$. This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density. In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)

@@

We can formally non-dimensionalize this system by choosing the general rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
\begin{align}
\begin{split}
\dot{F} &= \frac{1}{s}\left[\lambda_{max} F + \rho_{max}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
\dot{H} &= \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{max}\frac{R}{qk} H - \mu H\right],  \\
\dot{R} &= \frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right) F\right]\right].
\end{split}
\end{align}
If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
\begin{align}
\begin{split}
\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
\label{reduceddynamics}
\end{split}
\end{align}
where we have dropped the subscripts on $\lambda_{max}$ and $\rho_{max}$ for simplicity, and $\xi=C/k$, $\delta=Y_{H}kP_{H}/C$, and $\beta=Y_{H}k\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)/C$. The above equations represent the system of equations presented in the main text.
\\

@@

{\color{red} \section{Sensitivity to additional death terms}

It should be noted that our set of dynamics (Equations \ref{bigdynamics} and \ref{reduceddynamics}) could include a constant death term of the form $-d_{F}F$ and $-d_{H}H$ to represent death not directly linked to starvation. Adding terms of this form to our model would simply adjust the effective value of $\lambda$ and $\mu$, and we could rewrite Equation \ref{reduceddynamics} with $\lambda^{\prime}=\lambda-d$ and $\mu^{\prime}=\mu-d$. These substitutions would not alter the functional form of our model nor the steady-states and qualitative results, however the quantitive values could shift based on the size of $d$ relative to $\lambda$ and $\mu$.

Survivorship has a well-known functional form which changes systematically with size \cite{}. Typically survivorship is defined using the Gompertz curve 
\begin{equation}
F=F_{0}e^{\left(q_{0}/c\right)\left(1-e^{ct}\right)}
\label{gompertz}
\end{equation}
where the parameters have the following allometric dependencies on adult mass \cite{}.

We are interested in the specific death rate of the form $\cdot{F}=-dF$, and using the derivative of Equation \ref{gompertz} we find that $d=e^{ct}q_{0}$. Our model considers the average rates over a population and lifecycle and the average death rate is given by 
\begin{eqnarray}
\bar{d}&=&\frac{1}{t_{exp}}\int_{0}^{t_{exp}}e^{ct}q_{0} dt \\
&=&\frac{q_{0}\left(e^{c t_{exp}}-1\right)}{ct_{exp}}
\end{eqnarray}
where $t_{exp}$ is the expected lifespan following the allometry of $t_{exp}=$ \cite{}. Given the allometries above we have that
\begin{equation}
d=
\end{equation}
which scales roughly like $SCALING$. In figure \ref{} we compare the value of $d$ to the reproductive and starvation mortality rates . The values of $d$ are orders of magnitude smaller than these rates at all scales, and adding this non-starvation based death rate to our model does not shift our results within numerical confidence.

\begin{figure}
\centering
\includegraphics[width=0.48\textwidth]{fig_FPAllometric2.eps}
\caption{\small{ Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
  The data are from Damuth \citep{Damuth:1987kr} and have been converted to
  total population metabolism using the allometric relationships for
  metabolic rate (e.g. Refs.~\citep{West:2001bv,hou,moses2008rmo}).}\label{fig:equivalence}}
\end{figure}

}

{\color{red} \section{NSM and the energy equivalence hypothesis}

The energy equivalence hypothesis is based on the observation that if one assumes that the total metabolism of an ecosystem $B_{\rm tot}$ is equally partitioned between all species ($B_{i}$, the total metabolism of one species, is a constant), then the abundances should follow $N\left(M\right)B\left(M\right)=B_{i}$ implying that $N\left(M\right)\propto M^{-\eta}$, where $\eta$ is the metabolic scaling exponent \citep{allen2002,enquist1998}. As $\eta \approx 3/4$ this hypothesis is consistent with Damuth's law \citep{allen2002}, however the actually equivalence of energy usage of diverse species has not been measured at the population level for a variety of whole populations. Figure \ref{fig:equivalence} shows that $F^{*}B$ is nearly constant over this same range.

\begin{figure}
\centering
\includegraphics[width=0.48\textwidth]{fig_FPAllometric2.eps}
\caption{\small{ Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
  The data are from Damuth \citep{Damuth:1987kr} and have been converted to
  total population metabolism using the allometric relationships for
  metabolic rate (e.g. Refs.~\citep{West:2001bv,hou,moses2008rmo}).}\label{fig:equivalence}}
\end{figure}

}
